Andres Karjus

### 2018-06-05 index.Rmd
redesign

---
title: "Andres Karjus"
output: 
  html_document: 
    theme: spacelab
    toc: yes
    toc_depth: 2  
    toc_float:
      collapsed: no
---

<style>
body .main-container {
max-width: 1500px;
padding-left: 0px;
}

@@

#TOC {
  text-align: left;
  padding-left: 0px ;
  padding-right: 0px;
  left: 10px;
  top: 0;
  width: 100%;
}

</style>

# New things

- _Selection above the baseline: quantifying the advection effect in four domains of cumulative culture_ @ the [Applications in Cultural Evolution: Arts, Languages, Technologies](https://cultevol.ut.ee/): slides with a bunch of interactive plots that allow playing with the massive datasets [now available here](https://andreskarjus.github.io/cultevol_tartu_slides)

# Short academic bio

<IMG SRC="IMG_2657.jpg" ALIGN="right" BR CLEAR="left" style="margin-left:10px"; /> 
<div style="text-align: justify"> 
I am a PhD student in linguistics based at the [Centre for Language Evolution at the University of Edinburgh](http://www.lel.ed.ac.uk/cle/), working on evolutionary variant selection in language with [Kenny Smith](http://www.lel.ed.ac.uk/~kenny), [Simon Kirby](http://www.lel.ed.ac.uk/~simon) and [Richard A. Blythe](http://www2.ph.ed.ac.uk/~rblythe3/), utilizing massive corpora and natural language processing to hunt traces of selection and competition between elements of language. My research is supported by a Kristjan Jaak scholarship from the Archimedes Foundation of Estonia.

I am also affiliated with the University of Tartu as a (part-time) junior researcher in sociolinguistics at the [EKKAM](http://ekkam.ut.ee/en/) research group, specializing in data analysis and agent-based models.

# Research

## Publications

* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2018. Quantifying the dynamics of topical fluctuations in language. | [arXiv](https://arxiv.org/abs/1806.00699)
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2018. Topical advection as a baseline model for diachronic lexical dynamics. Proceedings of The Society for Computation in Linguistics. Volume 1. [extended abstract] | [open access](http://scholarworks.umass.edu/scil/) | <a href="karjus_topical_2018.bib" target="_blank">bib</a>
* Andres Karjus, Martin Ehala, [accepted]. Testing an agent based model of language choice on sociolinguistic survey data. Language Dynamics and Change.
*	Martin Haspelmath, Andres Karjus, 2017. Explaining asymmetries in number marking: Singulatives, pluratives and usage frequency. Linguistics, volume 55, issue 6. | [online version](https://www.degruyter.com/view/j/ling.2017.55.issue-6/ling-2017-0026/ling-2017-0026.xml)  | [preprint](https://www.academia.edu/25032500/Explaining_asymmetries_in_number_marking_Singulatives_pluratives_and_usage_frequency) | <a href="haspelmath_explaining_2017.bib" target="_blank">bib</a>
*	Andres Karjus, 2015. Through the Spyglass of Synchrony: Grammaticalization of the Exterior Space in the Eastern Circum-Baltic. In: Hilpert, Martin, Östman, Jan-Ola, Mertzlufft, Christine, Rießler, Michael, Duke, Janet (eds.), Advances in Nordic Linguistics. De Gruyter Mouton.
| [google books](https://books.google.ee/books?id=BPheCAAAQBAJ)
*	Andres Karjus (editor), 2013. Areal linguistics, Grammar and Contacts. Special issue of the Journal of Estonian and Finno-Ugric Linguistics, 4-2. Tartu: University of Tartu Press. 
| [open access](http://jeful.ut.ee/index.php/JEFUL/issue/view/2013.4.2)
* Petar Kehayov, Eva Saar, Miina Norvik, Andres Karjus, 2013. Hääbuva kesklüüdi murde jälgedel suvel 2012 [On the footsteps of vanishing Central Lude in the summer of 2012]. Yearbook of the Estonian Mother Tongue Society, Vol. 58. | [open access](http://www.kirj.ee/public/ESA/2012/esa_58-2012-58-101.indd.pdf)
*	Andres Karjus, 2012. Outdoors on the Shores of the Baltic: Gradience in the Grammaticalization of the Exterior-Region. Journal of Estonian and Finno-Ugric Linguistics 3-1, pp. 209-226.

## Conferences

@@

# Consulting, R workshops &bookings

[I am available for booking](http://writingcentre.ppls.ed.ac.uk/appointments/) as a written communication consultant via the Edinburgh University School of Psychology, Philosophy and Language Sciences Writing Centre - specializing in writing about and presenting (graphs and figures of) data and data analysis results (this service is available to students of the School only). I am also occasionally holding workshops on using R for visualizing data for the Writing Centre - the dates and booking information are available [here](http://writingcentre.ppls.ed.ac.uk/workshops/). Feel free to have a look at the workshop materials [here](https://andreskarjus.github.io/artofthefigure/).

I will be holding a full day session on data visualization using R at the [Scottish Graduate School of Social Science Summer School 2018](http://www.socsciscotland.ac.uk/skills_and_training/summer_school).

I have also organized similar things in the past: a workshop on corpus linguistics using R for the [Academia Salensis](http://academiasalensis.org) summer school of 2012, and a course on data analysis for digital humanities at the University of Tartu in 2015. I am also currently engaged in teaching stats and R to [Edinburgh Uni psychology](https://www.ed.ac.uk/ppls/psychology) masters students.

Feel free to get in touch if you are interested in inviting me to teach a workshop on anything between data science and statistics for the humanties and social sciences, R, corpus linguistics, natural language processing, digital humanities, etc.

# Other things
## Semi-academic

I currently co-organize the bimonthly [Edinburgh Language Lunch](http://groups.inf.ed.ac.uk/langlunch/). In the past, I was a co-organizer of the [Linguistics and English Language Postgraduate Conference](http://www.lel.ed.ac.uk/~pgc/) of 2017 in Edinburgh, and head organizer of the [Why Linguistics Conference](http://whylinguistics.ut.ee) (2015) and the [Graduate Conference on Areal Linguistics, Grammar and Contacts](http://arealling.ut.ee) (2012) in Tartu, Estonia. I also served as a student representative at various boards during my MA and MSC, and as a freshmen tutor during my BA.

## Non-academic

@@

# Contact

<a href="mailto:a.karjus--at--sms.ed.ac.uk">email</a>

>>>>

---
title: "Andres Karjus"
output: 
  html_document: 
    theme: spacelab
    toc: no
    toc_depth: 4  
    toc_float:
      collapsed: no
---

<style>

/* 
body .main-container {
max-width: 1500px;
padding-left: 0px;
}
 */

@@

#TOC {
  text-align: left;
  padding-left: 0px ;
  padding-right: 0px;
  left: 10px;
  top: 0;
  width: 100%;
}
</style>

# {.tabset .tabset-fade}

## Research

### New things

- _Selection above the baseline: quantifying the advection effect in four domains of cumulative culture_ @ the [Applications in Cultural Evolution: Arts, Languages, Technologies](https://cultevol.ut.ee/) conference: slides with a bunch of interactive plots that allow playing with the massive datasets [now available here](https://andreskarjus.github.io/cultevol_tartu_slides)

### What I do

<IMG SRC="IMG_2657.jpg" ALIGN="right" BR CLEAR="left" style="margin-left:10px; width: 60%; height: auto;"; /> 
<div style="text-align: justify"> 
I am a PhD student in linguistics based at the [Centre for Language Evolution at the University of Edinburgh](http://www.lel.ed.ac.uk/cle/), working on lexical dynamics and evolutionary variant selection in language with [Kenny Smith](http://www.lel.ed.ac.uk/~kenny), [Simon Kirby](http://www.lel.ed.ac.uk/~simon) and [Richard A. Blythe](http://www2.ph.ed.ac.uk/~rblythe3/), utilizing massive corpora and natural language processing to hunt traces of selection and competition between elements of language. My research is supported by a Kristjan Jaak scholarship from the Archimedes Foundation of Estonia. I am also affiliated with the University of Tartu as a (part-time) junior researcher in sociolinguistics at the [EKKAM](http://ekkam.ut.ee/en/) research group, specializing in data analysis and agent-based models. Besides research, I am also engaged in [teaching and consulting](https://andreskarjus.github.io/#teaching,_consulting_&_R workshops) data sciency things for the humanities.

### Research

#### Publications

* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2018. Quantifying the dynamics of topical fluctuations in language. | [arXiv preprint](https://arxiv.org/abs/1806.00699)
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2018. Topical advection as a baseline model for diachronic lexical dynamics. Proceedings of The Society for Computation in Linguistics. Volume 1. [extended abstract] | [open access](http://scholarworks.umass.edu/scil/) | <a href="karjus_topical_2018.bib" target="_blank">bib</a>
* Andres Karjus, Martin Ehala, [accepted]. Testing an agent based model of language choice on sociolinguistic survey data. Language Dynamics and Change.
*	Martin Haspelmath, Andres Karjus, 2017. Explaining asymmetries in number marking: Singulatives, pluratives and usage frequency. Linguistics, volume 55, issue 6. | [online version](https://www.degruyter.com/view/j/ling.2017.55.issue-6/ling-2017-0026/ling-2017-0026.xml)  | [preprint](https://www.academia.edu/25032500/Explaining_asymmetries_in_number_marking_Singulatives_pluratives_and_usage_frequency) | <a href="haspelmath_explaining_2017.bib" target="_blank">bib</a>
*	Andres Karjus, 2015. Through the Spyglass of Synchrony: Grammaticalization of the Exterior Space in the Eastern Circum-Baltic. In: Hilpert, Martin, Östman, Jan-Ola, Mertzlufft, Christine, Rießler, Michael, Duke, Janet (eds.), Advances in Nordic Linguistics. De Gruyter Mouton.
| [google books](https://books.google.ee/books?id=BPheCAAAQBAJ)
*	Andres Karjus (editor), 2013. Areal linguistics, Grammar and Contacts. Special issue of the Journal of Estonian and Finno-Ugric Linguistics, 4-2. Tartu: University of Tartu Press. 
| [open access](http://jeful.ut.ee/index.php/JEFUL/issue/view/2013.4.2)
* Petar Kehayov, Eva Saar, Miina Norvik, Andres Karjus, 2013. Hääbuva kesklüüdi murde jälgedel suvel 2012 [On the footsteps of vanishing Central Lude in the summer of 2012]. Yearbook of the Estonian Mother Tongue Society, Vol. 58. | [open access](http://www.kirj.ee/public/ESA/2012/esa_58-2012-58-101.indd.pdf)
*	Andres Karjus, 2012. Outdoors on the Shores of the Baltic: Gradience in the Grammaticalization of the Exterior-Region. Journal of Estonian and Finno-Ugric Linguistics 3-1, pp. 209-226.

#### Conferences

@@

## Teaching, consulting & R workshops

**Upcoming**
<div style="text-align: justify"> 
- I will be holding a full day session on data visualization using R at the [Scottish Graduate School of Social Science Summer School 2018](http://www.socsciscotland.ac.uk/skills_and_training/summer_school).

**Current**

- [I am available for booking](http://writingcentre.ppls.ed.ac.uk/appointments/) as a written communication consultant via the Edinburgh University School of Psychology, Philosophy and Language Sciences Writing Centre - specializing in writing about and presenting (graphs and figures of) data and data analysis results. This service is available to students of the School only. 
- I am also occasionally holding workshops on using R for visualizing data for the Writing Centre - the dates and booking information are available [here](http://writingcentre.ppls.ed.ac.uk/workshops/). Feel free to have a look at the workshop materials [here](https://andreskarjus.github.io/artofthefigure/).
- I am also currently engaged in teaching stats and R to [Edinburgh Uni psychology](https://www.ed.ac.uk/ppls/psychology) masters students.

**Past**

- aRt of the Figure data visualization workshop for the annual Edinburgh Linguistics Firbush retreat, spring 2018.
- Developed and co-taught a course on data analysis for digital humanities at the University of Tartu in the spring of 2016.
- Corpus linguistics using R course for the [Academia Salensis](http://academiasalensis.org) summer school of 2015.
- Assistant for courses on language technology and artificial intelligence, Department of Computer Science, University of Tartu, 2014-2016.
<br><br>

Feel free to get in touch if you are interested in inviting me to teach a workshop on anything related to data science and statistics for the humanties, data visualization, R, corpus linguistics, natural language processing, digital humanities, etc.
</div>

## Other things
### Semi-academic
<div style="text-align: justify"> 
I currently co-organize the bimonthly [Edinburgh Language Lunch](http://groups.inf.ed.ac.uk/langlunch/). In the past, I was a co-organizer of the [Linguistics and English Language Postgraduate Conference](http://www.lel.ed.ac.uk/~pgc/) of 2017 in Edinburgh, and head organizer of the [Why Linguistics Conference](http://whylinguistics.ut.ee) (2015) and the [Graduate Conference on Areal Linguistics, Grammar and Contacts](http://arealling.ut.ee) (2012) in Tartu, Estonia. I also served as a student representative at various boards during my MA and MSC.

### Non-academic

@@

### Before Scotland

## Contact

<a href="mailto:a.karjus--at--sms.ed.ac.uk"> @ </a> | <a href="https://twitter.com/AndresKarjus?ref_src=twsrc%5Etfw" class="twitter-follow-button" data-size="large" data-show-screen-name="false" data-show-count="false">@AndresKarjus</a><script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>
